Continous time Markov chain models; \\ extension, simulation and visualisation

### 2014-08-14 chapters/chapter01.tex
small changes

[PL: briefly introduce the field of molecular evolution, touching upon the essence of some basic evolutionary concepts (mutation<->substitution, selection, fitness), and how phylogenetics can be useful int this context. 
Describe the very basics of a tree (branches, nodes, taxa, rooted, outgroup), and provide some history on phylogenies and what they are used for. I would suggest to start by showing a figure combining Darwin's speciation tree in one panel, because that directly relates to the evolutionary and speciation process you need to bring up in the paragraph above, and the first estimated tree, which I believe is from Fitch in seventies science paper.
And then follow up on this with a brief account of phylogenetic inference -- bring on the classics, acknowledge some milestones: distance-based trees, parsimony inference, Joe and Likelihood inference (no formal stuff/math just yet), ...
This is also where you can describe the computational burden and that table.

%TODO: other problems

@@

Our group has been involved in research which aims at developing a full quantitative understanding of the processes that shape the epidemiology and the evolution of MEP-like viruses.
In the following section I present the problems on which my involvement in this fascinating area of research has been focused.

@@

[PL: Discuss variable size coalescent (demographic) models (which are the tree priors in BEAST)
Introduce phylogeography -- discrete ancestral state reconstruction is only one way of doing this (see Bloomquist, Lemey, Suchard)]

@@

[PL: How this all integrates in a flexible framework and publicly available software, one that adequately accounts for phylogenetic uncertainty by averaging over all plausible evolutionary histories for your data
If there is too little here to build a subsection, you can attach it to the previous subsection and call that 'Bayesian genealogical inference using BEAST'
]

>>>>

% 1 / 1.5 page, history and shit, old vs new tree plots (estimated and not), distance / parsimony, ML, Bayesian...
[PL: briefly introduce the field of molecular evolution, touching upon the essence of some basic evolutionary concepts (mutation<->substitution, selection, fitness), and how phylogenetics can be useful int this context. 
Describe the very basics of a tree (branches, nodes, taxa, rooted, outgroup), and provide some history on phylogenies and what they are used for. I would suggest to start by showing a figure combining Darwin's speciation tree in one panel, because that directly relates to the evolutionary and speciation process you need to bring up in the paragraph above, and the first estimated tree, which I believe is from Fitch in seventies science paper.
And then follow up on this with a brief account of phylogenetic inference -- bring on the classics, acknowledge some milestones: distance-based trees, parsimony inference, Joe and Likelihood inference (no formal stuff/math just yet), ...
This is also where you can describe the computational burden and that table.

@@

Our group has been involved in research which aims at developing a full quantitative understanding of the processes that shape the epidemiology and the evolution of MEP-like viruses.
In the following section I present the problems on which my involvment in this fascinating area of research has been focused.

@@

Discuss variable size coalescent (demographic) models priors (which are
Introduce phylogeography -- discete ancestral state reconstruction is only one way of doing this (see Bloomquist, Lemey, Suchard)

@@

How this all integrates in a flexible framework and publicly available software

### 2014-08-14 chapters/chapter05.tex
small changes

In the GY94-F$1\times4$ model codon frequencies are expected from the four nucleotide frequencies, with frequencies being pulled from all three codon positions, i.e. $\pi_{n_{1}}=\pi_{n_{3}}=\pi_{n_{3}}$ for all four nucleotides,.
Thi smodel leads to 3 free parameters that need to be estimated from the data.
If the frequencies are parametrized according to three sets of nucleotide frequencies for the three codon positions, resulting in nine free parameters, the model is called the GY94-F$3\times4$.
Finally in the GY94-F61 model every codon triplet has it's own frequncy parameter, with all parameters summing to one, resulting in 60 free parameter that need to be estimated.

@@

Their \emph{neutral model} assumes a category for sites where the nonsynonymous mutations are neutral ($\omega=1$) and a category for sites which are conserved ($\omega=0$).
\emph{Positive selection model} adds an extra category of positively selected sites ($\omega>1$).

>>>>

In the GY94-F$1\times4$ model codon frequencies are expected from the four nucleotide frequencies, with frequencies being pulled from all three codon positions, i.e. $\pi_{n_{1}}=\pi_{n_{3}}=\pi_{n_{3}}$ for all four nucleotides,.
This model leads to 3 free parameters that need to be estimated from the data.
If the frequencies are parametrized according to three sets of nucleotide frequencies for the three codon positions, resulting in nine free parameters, the model is called the GY94-F$3\times4$.
Finally in the GY94-F61 model every codon triplet has it's own frequency parameter, with all parameters summing to one, resulting in 60 free parameter that need to be estimated.

@@

Their \emph{neutral model} assumes a category for sites where the nonsynonymous mutations are neutral ($\omega_1=1$) and a category for sites which are conserved ($\omega_2=0$).
\emph{Positive selection model} adds an extra category of positively selected sites ($\omega_3>1$).
